### 2011-10-18 doc/manuscript_plos.tex
I have to keep revisioning of the manuscript.

@@

Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  We aims at which parts of the genomic
organization and human pathogenic virulence factors of SDE could have originated
in those of SPY, or vice versa.  Our genome-wide survey of gene transfer
indicated that gene transfer via homologous recombination occurred more often
for the net direction from SPY to SDE.  We found evidence that virulence genes
transferred more often via non-recombining fashion without replacing parts of
destination genomes than did non-virulence genes.  A genome-wide resource of
recombining and non-recombining gene transfer for the study of the SPY and SDE
genomes is provided so that research community in the pathogenic SPY and SDE
bacteria could take advantage of it.

@@

Homologous recombining gene transfer could be studied more readily using the
core genomes of individual bacteria while horizontal gene transfer might be more
easily detected using the pan genomes. Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segments between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.  Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene duplication, loss, and
transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  We noted that the first
approach could be used to infer transfer of genomic segments due to homologous
recombination, and the second could be useful in studying horizontal gene
transfer.

In short, the complete genomes of the two species of SPY and SDE presented us a
unique opportunity to study gene transfer over relatively recent time scales.
We took advantage of the model-based approach for inferring homologous
recombination, and the parsimony-based one for detecting gene duplication, loss,
and horizontal transfer. Virulence genes were found to be more associated
non-recombining horizontal gene transfer than non-virulence genes.  We compared
both of the results in recombining gene transfer events to show that the
parsimony-based method inferred recombining gene transfer events in biological
functional categories similar to those inferred with the model-based approach.
This was the first attempt as far as we know of systematically analyzing SPY and
SDE in genome-wide scale to detect gene transfer events not only in the core
genome but also in the pan genome, which made the analysis complete in terms of
the amount of used genome data. Our analysis of core genomes also showed for the
first time that the net directionality of recombining gene transfer from SPY to
SDE was stronger than the reverse direction.

@@

@@

We tried in vain to quantify differential horizontal gene transfer between SPY
and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
the clade-crossing transfers, we did not see any significant enrichment for a
specific direction, as opposed to the result for recombining transfers from the
model-based analysis.  SDE1 and SDE2 shared the most events (26 and 34) compared
to other sample pairs.  SDD appeared to acquire many genes from the ancestral
SDE branch.  Transfer events between the SPY and SDE clades were inferred in
smaller numbers (<6) compared to those within each clade. We were puzzled by
this finding because virulence genes were found to be associated with horizontal
gene transfer, and yet we could not show net directionality of horizontal gene
transfer between SPY and SDE.  However, we inferred relatively more recombining
gene transfer events from the SPY branch to SDE branch even in the
parsimony-based approach (Figure S\ref{fig:mowgli-recomb-heatmap}).

@@

About 10\% of the genes comprised of virulence facters, about half of which were
found in the core genome of the five genome alignment \citep{Suzuki2011}.  Using
customized UCSC genome browser tracks available at
\url{http://strep-genome.bscb.cornell.edu} we located virulence genes with
notable posterior recombination probability.  These genes were by no means a
comprehensive list of virulence factors that might have been transferred via
homologous recombination.  We wished to showcase virulence genes using the
result of our strudy of gene transfer.

@@

A full model-based approach would be desirable in analyzing data for testing
relevant hypotheses more systematically. Model complexity might not allow easy
access to the implementation of a model. Our study involved a model-based
approach and a parsimony-based one for understanding evolution in the core
genomes of several bacteria individuals. Because of inability of the model-based
method for studying the pan genomes we had to resort to the parsimony-based
method. However, the parsimony-based method was also applied to the core
genomes. We wished to compare the two methods in their results of the core
genome analyses.  Because of two independent analyses on the same data set, this
would allow us to be more confident of the results if the two analyses performed
similarly Generally, phylogenetic methods of detecting gene transfers often take
advantage of discrepancy between a reference species tree and gene trees. The
two approaches that we employed in this study were no exception. We found a few
similar results in both methods; firstly, functional categories that were
associated with recombining gene transfer were similar (Tables
S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Secondly,
either of them did not show association of virulence genes with recombining gene
transfer.  To more directly compare the methods we used the recombination
intensity and the number of recombining gene transfer events for each gene.  We
found a weak positive correlation of 0.39 with a very small significant P-value
of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
zero.  (Figure S\ref{fig:cmpcomowgli}).  Interestingly, the correlation (0.03
with P-value of 0.18) disappeared in a comparison between recombination
intensity and number of horizontal gene transfer events (Figure
S\ref{fig:cmpcomowglihgt}).

@@

The process of gene transfer may or may not move whole genes from one species to
another \citep[e.g.,][]{Chan2009}. We observed that units of recombining gene
transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
cases, parts of a gene were inferred to have been transferred, for instance,
transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
gene. Recombinant tracts often were shown to span two genes.  In other cases,
the inferred transfer units spaned a set of as much as eight adjacent genes:
loci from SpyM3\_1518 to SpyM3\_1525.

\subsection*{Transporters, transferases, aminoacyl tRNA synthetases, and
transposases}

Molecular evolutionary forces could act on genetic materials of organisms, and
changes to the genotypes might lead to some phenotypes that could change
biological functions of the species. So, the bacterial gene transfer process as
an evolutionary force could be associated with some biological functions. We
attempted to find which biological functions were associated with recombining or
horizontal gene transfers.  Several functional categories that were found to be
associated with recombining gene transfer were related with either
membrane-bound molecules or aminoacyl tRNA synthetases (Table
\ref{tab:functional}).  However, functional categories associated with
horizontal gene transfer were different from those from recombining gene
transfer analysis (Table \ref{tab:go-events}).  The top ranked functional
category associated with horizontal gene transfer was related with transposases.
These differences in the profile of functional categories between the two
processes were interesting to us. An approach of dicriminating the two gene
transfer processes would result in the differences.

Because bacteria as a unicellular organism should be able to respond to changes
to environment, some evolutionary pressures must have been acting on genes
encoding membrane-bound molecules. Recombining gene transfer process seemed to
be more associated with biological functions that had been critical of sensing
queues of bacterial environmental changes.  Sugar phosphotransferase classes
were ranked top in the list; phosphrylation could be important in signal
transduction to recognize environmental changes.  The transport functional
categories of carbohydrate transport and sodium ion transport also belonged to
genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
lipid biosyntheses that must be involved in producing membrane-bound molecules.
Horizontal gene transfer process would allow bacteria to acquire foreign genes
without the requirement of homologous regions in the bacteria.  Transposases
would help foreign genes be excised from a donor genome and integrated into a
recipient genome.

In addition to these biological categories, we also found several aminoacyl tRNA
synthetases (GO Term translation GO:0006412) in association with recombining
gene transfer.  In the parsimony-based approach the term translation was top
ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
transfer in bacterial species \citep{Woese2000}.  Some of the functional
categories were unexpected; functional categories concerning ribosomes were
found to be associated with recombining gene transfer.  Because genes related
with ribosomes were considered to be housekeeping genes, gene transfer in which
would be rare, it was unexpected that ribosome related genes were found to be
associated with recombining gene transfer.  Because it was a bizarre result, we
investigated the association by considering different recombinant edge types. We
noted that two different classes of recombinant edge types: one that changes
species tree topologies, and another that does not. Because recombinant edges
that did not change species tree topology might be difficult to infer reliably,
considering those recombinant edges in computing recombination intensity might
lead to dubious results.  Ribosome related functional categories were found to
be top ranked if we took those recombinant edges that did not change species
tree topology (Tables S\ref{tab:functional-notopology} and
S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
functional categories were not top ranked.  Therefore, high recombination
intensity in ribosome related genes were likely to be attributed to
topology-not-changing recombinant edges. We would be cautious of recombining
transfer of genes related with ribosome related functions.

@@

for each recombinant edge type to assess the effect of posterior recombinant
trees samples with respect to that of prior recombinant tree samples. The ten
\textit{a priori} recombinant trees from Simulation 2 would have balanced
recombining gene transfer events between SPY and SDE with respect to the prior
expected number of recombinant edges. The posterior recombinant trees used in
Simulation 3 would have unbalanced recombining gene transfer between the two
species branches as we observed in the real data analysis. For significantly
strong gene flows such as ones from SPY to SDE a solid-line interval with a
closed circle would be different from its count part of the dotted-line
interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
SDE was stronger enough to be nearer to the horizontal line at the zero on the
Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
too weak.  This comparison supported the biased directionality of recombining
gene transfer from SPY to SDE.

>>>>

\section*{Things To Do for submitting this to PLoS Pathogens}
\begin{enumerate}
\item A paragraph for briefing the methods of ClonalOrigin and Mowgli before
section \textbf{Results}.
\item We can add at least 50 words in section \textbf{Abstract}.
\item Figures, tables, and supporting information need to be ready.
\item Publish source codes at public domain.
\end{enumerate}

\section*{A list our major and minor findings or observations}
\begin{enumerate}
\item We showed that recombining gene transfer from SPY to SDE was strong
enough.
\item A few evidences for the net directionality from SPY to SDE:
time-constrained gene transfer between SPY to SDE also showed the directionality
from SPY to SDE.
\item We showed that virulence genes were associated with horizontal gene
transfer. 
\item We found that recombining gene tranfer processes are involved in
membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
\item We found that horizontal gene transfer processes are involved in
transposases.
\item We provided evidence that recombination process affected more sites along
the core genome of SPY and SDE than mutation process.
\item We found a recombination hotspot at 200 kb from the origin of replication
along the genome of SDE1.
\item We showed that species-tree-topology-changing recombinant edges were more
reliably recovered than species-tree-topology-not-changing edges.
\item We showed that recombination intensity might be useful in finding genes
under high recombining gene transfer.
\item We found that SPY and SDE are more closely related than SDD and SDE in
about 10\% of the core genome.
\item Recombination intensity and number of recombining gene transfer events
were positively correlated.
\item We found reduction of numbers of genes after divergence of SDE and SPY.
\item We confirmed the species tree that relates the five genomes.
\item We listed several virulence genes associated with recombining and
horizontal gene transfer.
\end{enumerate}

\section*{Suggestions about the manuscript}
\begin{enumerate}
\item Ilan thought that a species tree must be shown at the outset of the
section \textbf{Results}.
\item Ilan suggested a different order of four parts in \textbf{Results}:
preliminary, funtional category and virulence gene analyses, gene gain and loss,
and directionality.
\end{enumerate}

@@

Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  We aims at which parts of the genomic
organization and human pathogenic virulence factors of SDE could have originated
in those of SPY, or vice versa.  Two mechanisms of gene transfer in bacteria
were known; recipient bacteria of transferred genes can substitute transferred
genes for homologous parts of their genomes (Recombining Gene Transfer), or
integrate transferred genes into their genomes without replacement (Horizontal
Gene Transfer).  We discriminated the two processes by using a model-based
approach for studying recombining gene transfer, and using a parsimony-based
approach for both transfer processes.  Our genome-wide survey of gene transfer
indicated that gene transfer via homologous recombination occurred more often
for the net direction from SPY to SDE.  We also found evidence that virulence
genes transferred more often via non-recombining fashion without replacing parts
of destination genomes than did non-virulence genes.  
A genome-wide resource of
recombining and non-recombining gene transfer for the study of the SPY and SDE
genomes is provided so that research community in the pathogenic SPY and SDE
bacteria could take advantage of it.

@@

Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segments between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.   
% Descriptions of ClonalOrigin model.
Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene duplication, loss, and
transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  
% Descriptions of Mowgli analsysis.
We noted that the first
approach could be used to infer transfer of genomic segments due to homologous
recombination, and the second could be useful in studying horizontal gene
transfer.
% Homologous recombining gene transfer could be studied more readily using the
% core genomes of individual bacteria while horizontal gene transfer might be more
% easily detected using the pan genomes.

In short, the complete genomes of the two species of SPY and SDE presented us a
unique opportunity to study gene transfer over relatively recent time scales.
We took advantage of the model-based approach for inferring homologous
recombination, and the parsimony-based one for detecting gene duplication, loss,
and horizontal transfer. Virulence genes were found to be more associated
non-recombining horizontal gene transfer than non-virulence genes.  We compared
both of the results in recombining gene transfer events to show that the
parsimony-based method inferred recombining gene transfer events in biological
functional categories similar to those inferred with the model-based approach.
This was the first attempt as far as we know of systematically analyzing SPY and
SDE in genome-wide scale to detect gene transfer events not only in the core
genome but also in the pan genome, which made the analysis complete in terms of
the amount of used genome data. Our analysis of the core genomes also showed for the
first time that the net directionality of recombining gene transfer from SPY to
SDE was stronger at the genome scale.

@@

\subsection*{Distinguishing two gene transfer processes shows different functional
categories associations}

Bacterial gene transfer processes as an evolutionary force would affect more
frequently in some biological functions than in others.  In our attempt of
finding which biological functions were associated with recombining or
horizontal gene transfers, we devised a measure of recombination intensity for
genes by using frequency of recombinant edges that affect those genes, and used
frequency of horizontal gene transfer events (see section \textbf{Materials and
Methods}).  Table \ref{tab:functional} shows a list of function categories
that were found to be associated with recombining gene transfers. We noted three
biological functions in the list: membrane-bound molecules, aminoacyl tRNA
synthetases, and ribosomes.

Firstly, we have sugar phosphotransferases ranked
top in the list; phosphrylation could be important in signal transduction to
recognize environmental changes.  Transport functional categories of
carbohydrate transport and sodium ion transport would belong to genes encoding
membrane-bound molecules.  Also highly ranked were fatty acid and lipid
biosyntheses that must be involved in producing membrane-bound molecules. 
Because bacteria as a unicellular organism should be able to respond to changes
to environment, some evolutionary pressures must have been acting on genes
encoding membrane-bound molecules. Therefoure, recombining gene transfer process
appeared to be more associated with biological functions that had been critical
of sensing queues of bacterial environmental changes.

Secondly, we found several aminoacyl tRNA synthetases (GO Term translation
GO:0006412) in association with recombining gene transfer.  In the
parsimony-based approach the term translation was also top ranked (Table
S\ref{tab:go-events-recombining}).  Because they had been a focus of gene
transfer in bacterial species \citep{Woese2000}, we could confirmed that the
recombination intensity and requency of recombining gene transfer events would
be useful in finding genes associated with recombining gene transfer.

Thirdly, functional categories concerning ribosomes were unexpectedly
found to be associated with recombining gene transfer.  Because genes related
with ribosomes were considered to be housekeeping genes, gene transfer in which
would be rare, it was unexpected that ribosome related genes were found to be
associated with recombining gene transfer.  Because it was a bizarre result, we
investigated the association by considering different recombinant edge types. We
noted that two different classes of recombinant edge types: one that changes
species tree topologies, and another that does not
(see section \textbf{Materials and Methods}).
Because recombinant edges
that did not change species tree topology might be difficult to infer reliably,
considering those recombinant edges in computing recombination intensity might
lead to dubious results.  Ribosome related functional categories were found to
be top ranked if we took those recombinant edges that did not change species
tree topology (Tables S\ref{tab:functional-notopology} and
S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
functional categories were not top ranked.  Therefore, high recombination
intensity in ribosome related genes were likely to be attributed to
topology-not-changing recombinant edges. We would be cautious of recombining
transfer of genes related with ribosome related functions.

Table \ref{tab:go-events} shows a list of functional categories associated with
horizontal gene transfer, which were different from those from recombining gene
transfer analysis.  Ranked top in the list was transposases, which were known to
help foreign genes be excised from a donor genome and integrated into a
recipient genome. The biological role of transposases was concordant with
horizontal gene transfer process because the process involved in transposases
would allow bacteria to acquire foreign genes without the requirement of
homologous regions in the bacteria.  
These differences in the profile of functional categories association of the two
gene transfer processes were because of the approach of dicriminating the two
gene transfer processes.

\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}

We show a net directionality of recombining gene transfers from SPY to SDE.

Describe our arguments of the net directionality. 
Number of recombinant edges.
Confidence about the numbers of recombinant edges.
Not all of genes experience recombining gene transfers from SPY to SDE.
Gene transfer counts from the parsimony-based approach; the power of the
parsimony-based approach may be limited.
Is this number of recombinant edges good enough for the argument?

Reflection of the previous analyses, 
heat maps of real data analysis, 
and supporting simulations.

@@

We tried in vain to quantify differential horizontal gene transfer between SPY
and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
the clade-crossing transfers, we did not see any significant enrichment for a
specific direction, as opposed to the result for recombining transfers from the
model-based analysis.  SDE1 and SDE2 shared the most events (26 and 34) compared
to other sample pairs.  SDD appeared to acquire many genes from the ancestral
SDE branch.  Transfer events between the SPY and SDE clades were inferred in
smaller numbers ($<6$) compared to those within each clade. We were puzzled by
this finding because virulence genes were found to be associated with horizontal
gene transfer, and yet we could not show net directionality of horizontal gene
transfer between SPY and SDE.  However, we inferred relatively more recombining
gene transfer events from the SPY branch to SDE branch even in the
parsimony-based approach (Figure S\ref{fig:mowgli-recomb-heatmap}).

@@

About 10\% of the genes comprised of virulence facters, about half of which were
found in the core genome of the five genome alignment \citep{Suzuki2011}.  Using
customized UCSC genome browser tracks available at
\url{http://strep-genome.bscb.cornell.edu} we located virulence genes with
notable posterior recombination probability.  These genes were by no means a
comprehensive list of virulence factors that might have been transferred via
homologous recombination.  We wished to showcase virulence genes using the
result of our study of gene transfer.

@@

A full model-based approach would be desirable in analyzing data for testing
relevant hypotheses more systematically. Model complexity might not allow easy
access to the implementation of a model. Our study involved a model-based
approach and a parsimony-based one for understanding evolution in the core
genomes of several bacteria individuals. Because of inability of the model-based
method for studying the pan genomes we had to resort to the parsimony-based
method. However, the parsimony-based method was also applied to the core
genomes. We wished to compare the two methods in their results of the core
genome analyses.  Because of two independent analyses on the same data set, this
would allow us to be more confident of the results if the two analyses performed
similarly. Generally, phylogenetic methods of detecting gene transfers often take
advantage of discrepancy between a reference species tree and gene trees. The
two approaches that we employed in this study were no exception. We found a few
similar results in both methods; firstly, functional categories that were
associated with recombining gene transfer were similar (Tables
S\ref{tab:functional-all} and S\ref{tab:go-events-recombining}).  Secondly,
either of them did not show association of virulence genes with recombining gene
transfer.  To more directly compare the methods we used the recombination
intensity and the number of recombining gene transfer events for each gene.  We
found a weak positive correlation of 0.39 with a significant P-value
of $2.2\times10^{-16}$ for testing a null hypothesis of the correlation being
zero.  (Figure S\ref{fig:cmpcomowgli}).  Interestingly, the correlation (0.03
with P-value of 0.18) disappeared in a comparison between recombination
intensity and number of horizontal gene transfer events (Figure
S\ref{fig:cmpcomowglihgt}).

@@

The process of gene transfer may or may not move whole genes from one species to
another \citep{Chan2009}. We observed that units of recombining gene
transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
cases, parts of a gene were inferred to have been transferred, for instance,
transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
gene. Recombinant tracts were often shown to span two genes.  In other cases,
the inferred transfer units spaned a set of as much as eight adjacent genes:
loci from SpyM3\_1518 to SpyM3\_1525.

@@

For each recombinant edge type to assess the effect of posterior recombinant
trees samples with respect to that of prior recombinant tree samples. The ten
\textit{a priori} recombinant trees from Simulation 2 would have balanced
recombining gene transfer events between SPY and SDE with respect to the prior
expected number of recombinant edges. The posterior recombinant trees used in
Simulation 3 would have unbalanced recombining gene transfer between the two
species branches as we observed in the real data analysis. For significantly
strong gene flows such as ones from SPY to SDE a solid-line interval with a
closed circle would be different from its count part of the dotted-line
interval.  Figure \ref{fig:h3} showed that the recombining gene flow from SPY to
SDE was stronger enough to be nearer to the horizontal line at the zero on the
Y-axis. However, the recombining gene transfer from SDE to SPY appeared to be
too weak.  This comparison supported the biased directionality of recombining
gene transfer from SPY to SDE.
